---
title: "Modelling project"
author: "Himal Shrestha"
date: "28/06/2021"
output: 
  md_document:
    variant: markdown_github
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/19226876/OneDrive - LA TROBE UNIVERSITY/PhD Year II/Workshops/210623_data driven modelling")
```

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(session)
```

```{r}
restore.session(file = "data_modelling.RSession")
```

```{r}
save.session(file = "data_modelling.RSession")
```

### Load data

```{r, warning=FALSE, echo=FALSE}
births <- read.csv("data/births.csv")
cases <- read.csv("data/chicken_pox_cases.csv")
population <- read.csv("data/population.csv")
```

```{r}
qplot(year, births, data = births)

ggplot(cases) +
  geom_line(aes(year, cases)) + xlim(1950, 1960)

qplot(year, population, data = population)
```

Parameters to be estimated:

-   $\alpha$ - birth rate

-   $\mu$ - mortality rate

-   $b$ -

-   $\gamma$ - recovery rate

-   $\beta_0$ - average infection rate throughout the year

-   $\epsilon$ - strength of seasonal forcing function

-   $S_0$ - number of susceptible individuals at the start of 1950

## Fitting the ODE model

### $\alpha$ parameter

```{r}
births_50s <- births %>% filter(year >= 1950 & year < 1960)
birth <- sum(births_50s$births)/(max(births_50s$year) - min(births_50s$year))
```

### $\mu$ parameter

```{r}
population_50s <- population %>% filter(year >= 1950 & year < 1960)
pop_diff <- (max(population_50s$population) - min(population$population))/(max(population_50s$year) - min(population_50s$year))
death_rate <- pop_diff - birth
```

### $b$ parameter

```{r}
cases_50s <- cases %>% filter(year >= 1950 & year < 1960)
plot(cases_50s$year, cases_50s$cases)
```

```{r}
t <- cases_50s$year
y <- cases_50s$cases

# ssp <- spectrum(y)
# per <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
# reslm <- lm(y ~ sin(2*pi/365*t) +cos(2*pi/365*t))
# summary(reslm)
# 
# rg <- diff(range(y))
# plot(y~t,ylim=c(min(y)-0.1*rg,max(y)+0.1*rg))
# lines(fitted(reslm)~t,col=4,lty=2)

res <- nls(y ~ e*sin((2*pi/365)*t + b), start = list(e = 600, b = .5))
co <- coef(res)

fit <- function(e, x, b) {e*sin((2*pi/1)*x + b)}

pred_cases <- fit(e = co["e"], x = t, b = co["b"])
plot(t, y)
curve(fit(x, e = co["e"], b = co["b"]), col = "blue", add = T)
```
## SIR
Running simple SIR model in R
```{r}
library(deSolve)
```

1. Define R function for the general system of equations
```{r}
SEI_seasonal = function(time, state, parameters){
  with(as.list(c(state, parameters)),{
    beta <- beta_0 * (1 + epsilon * sin((2 * pi * time/1) +  b))
    
    # defining the equations
    dS <- alpha * (S + I + R) - mu * S - beta * S * I
    dI <- beta * S * I - (gamma + mu) * I
    dR <- gamma * I - mu * R
    res = c(dS, dI, dR, beta)
    # return the list of gradients
    list(res)
  })
}

```

2. Specify the time points at which we want the integrator to save the state of the system
3. Provide values for the parameters
4. Give initial values for the all state variables
```{r}
times  = seq(1950, 2010, by = 1/100)

SEI_seasonal_pars <- c(
  alpha = 139200.9,
  mu = 255449.1,
  beta_0 = 900000,
  epsilon = 979.6160047,
  b = 0.5636866,
  gamma = (1/2)
)

S_0 <- 4597542 - 385 - 255449.1 + 139200.9

init <- c(S = S_0, I = 385, R = 1000, beta = )
```

5. Invoke R function to do the integration
```{r}
out <- ode(y = init, times = times, func = SEI_seasonal, parms = SEI_seasonal_pars)
out=as.data.frame(out) 
head(round(out, 3))
```
```{r}
# Plot the output ---------------------------------------------------------
tidy_out <- out %>% pivot_longer(!time, names_to = "variable", values_to = "values") %>% filter(variable == "I")

ggplot(tidy_out) +
  geom_line(mapping = aes(x = time, y = values, color = variable))
```

```{r}
plot(x= out$time , y=out$S, ylab="Fraction", xlab= "Time", type="l")
lines(x=out$time, y=out$I, col="red") 
lines(x=out$time, y=out$R, col="green")
```
